\makeatletter
\def\input@path{{MST/}}
\makeatother
\documentclass{iopjournal}
\usepackage{amsmath,amsfonts,amssymb,bm}
\usepackage{mathtools}
\usepackage{algorithmic}
\usepackage{algorithm}
\usepackage{array}
\allowdisplaybreaks
\usepackage{subfig}
\usepackage{textcomp}
\usepackage{url}
\usepackage{verbatim}
\usepackage{cite}
\usepackage{booktabs}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}

% --- Math shortcuts ---
\newcommand{\R}{\mathbb{R}}
\newcommand{\E}{\mathbb{E}}
\newcommand{\norm}[1]{\left\lVert#1\right\rVert}
\newcommand{\diag}{\operatorname{diag}}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\newcommand{\IEEEPARstart}[2]{#1#2}
\newcommand{\IEEEmembership}[1]{#1}

\begin{document}

\articletype{Paper}

\title{Hierarchical Tactical Corridor Planning and Control\\
for Wheel-to-Wheel Autonomous Racing}

\author{Shuyi Gao, Defu Lin and Ren Jin}

\affil{School of Aerospace Engineering, Beijing Institute of Technology, Beijing, China}

\keywords{Autonomous racing, model predictive contouring control, second-order cone
programming, reinforcement learning, Frenet frame, friction circle,
game-theoretic motion planning.}

\begin{abstract}
Wheel-to-wheel autonomous racing couples strategic interaction, real-time
trajectory optimization, and friction-limit control.  Solving these decisions
in a single monolithic game or nonlinear program is computationally fragile at
Formula-class speeds, whereas end-to-end neural control is difficult to audit
on a safety-critical vehicle.  This paper presents a hierarchical architecture
that restricts learning to tactical corridor synthesis and retains
optimization-based planning and control downstream.  A map-conditioned
Truncated Quantile Critics policy and a hysteretic finite-state machine select
an overtake side, velocity cap, and time-varying Frenet corridor.  A sparse
second-order cone program computes a curvature-regularized lateral path within
that corridor, a forward--backward sweep imposes a super-elliptic GG
friction envelope, and a nonlinear model predictive contouring controller
tracks the resulting reference with a terminal progress reward and a soft
rear-axle friction constraint.  Closed-loop simulations on the Dallara EAV24
at Yas Marina demonstrate two successful fast-closure overtakes, with pass
times of $1.13\,\mathrm{s}$ and $8.63\,\mathrm{s}$ after entering
\textsc{Overtake}, peak speed up to $73.10\,\mathrm{m/s}$, and lateral RMS
tracking errors of $0.02$ and $0.05\,\mathrm{m}$.  The results support the
central premise that
learning should shape the feasible set, while model-based layers enforce
vehicle dynamics and handling-limit execution.
\end{abstract}

% =====================================================================
\section{Introduction}
% =====================================================================
\IEEEPARstart{A}{utonomous} racing at the vehicle handling limit is among
the most demanding closed-loop control problems in robotics.  A Formula-class
car must decide \emph{whether} to overtake, \emph{which side} to attempt,
\emph{where} to place the resulting trajectory within the track boundaries,
and \emph{how} to execute it at the friction limit---all while an adversarial
opponent simultaneously optimizes its own plan.  These four decisions span
discrete game theory, continuous trajectory optimization, and high-bandwidth
nonlinear control, operating across timescales that differ by three orders of
magnitude.  The Abu Dhabi Autonomous Racing League (A2RL), with its
$5.28\,\mathrm{km}$ Yas Marina Formula~1 circuit and the $700\,\mathrm{kg}$
Dallara EAV24 platform, provides a uniquely demanding testbed for this
hierarchical challenge~\cite{betz2023autonomous,wischnewski2022indy,jung2023iac}.

Classical autonomous driving architectures decompose the problem into
perception, behavior planning, motion planning, and
control~\cite{paden2016survey,gonzalez2016review}.  The behavior planner
selects a discrete maneuver---follow, change lane, yield---which the motion
planner translates into a continuous trajectory~\cite{werling2010frenet,ziegler2014baidu}.
In racing, however, the motion planner has traditionally been cast as a
single-vehicle, offline lap-time optimization
problem~\cite{christ2021,heilmeier2020,kapania2016,lipp2014,verschueren2014},
producing a fixed raceline tracked by an MPC or MPCC
controller~\cite{liniger2015,liniger2019rss,kabzan2019gpmpc,rosolia2020lmpc,brunner2017}.
Full-stack racing platforms developed for the Indy Autonomous Challenge and
A2RL document this paradigm in
detail~\cite{kabzan2020amzracing,betz2023autonomous,jung2023iac,raji2024kth}.
The fundamental limitation is that a pre-computed raceline cannot adapt to
an opponent who is simultaneously optimizing their own trajectory: an
adversary that defends the apex creates a geometric deadlock that a
single-vehicle planner cannot resolve without a higher-level instruction
about which side to attempt the pass and when to commit.

Recognising that wheel-to-wheel racing is inherently a non-cooperative
dynamic game, several groups have proposed game-theoretic formulations of the
tactical layer.  Wang \emph{et al.}~\cite{wang2021iterative} cast
head-to-head racing as an iterative best-response game with local Nash
convergence guarantees.  Iterative linear-quadratic game
solvers~\cite{fridovich2020ilqgames,laine2023} compute approximate
generalized feedback Nash equilibria at near-real-time rates.  Real-time
game-theoretic planners~\cite{spica2020} and viability-theoretic
approaches~\cite{liniger2020stackelberg,liniger2019rss} provide formal safety
certificates, while sampling-based MPPI~\cite{williams2018best} and Bayesian
intent inference~\cite{jia2023bayesianrace,sunberg2017pomdp} address
opponent-model uncertainty.  Despite this progress, pure game-theoretic
methods face four practical obstacles in Formula-class deployment:
(i)~exact Nash equilibria of continuous-time differential games are
intractable, and approximate solvers~\cite{fridovich2020ilqgames,laine2023}
may not converge within a single planning cycle at circuit speeds;
(ii)~the Nash solution is fragile to misspecification of the opponent's cost
function; (iii)~the discrete decision to switch from shadowing to overtaking
is not naturally captured by smooth equilibrium computations; and
(iv)~hand-tuned weights and horizon lengths must be recalibrated for each
track and pace configuration~\cite{betz2022software,betz2022survey}.

In parallel, data-driven methods have emerged as an alternative.  Sony's
GT~Sophy~\cite{wurman2022gtsophy} and subsequent RL-for-racing
work~\cite{fuchs2021superhuman,song2021autonomous,kaufmann2023champion}
demonstrated that end-to-end neural policies can achieve superhuman
performance in simulation.  Model-based RL~\cite{brunnbauer2022learning,schwarting2021},
safe RL~\cite{evans2023safelearning,brunke2022safe,he2023ralrace},
and learning-from-demonstration~\cite{remonda2024,rosolia2018sample}
have further expanded the data-driven toolkit.  Lin \emph{et
al.}~\cite{lin2024racellm} compared optimal control and RL at the handling
limit, finding that RL excels in multi-agent settings but underperforms in
single-vehicle lap time.  However, deploying end-to-end neural controllers
on a physical Formula car raises four serious concerns: (i)~the EAV24's
nonlinear tire model, aerodynamic downforce map, and differential torque
vectoring are not faithfully reproduced in open simulators; (ii)~a neural
network that outputs steering torque cannot be audited or certified by race
engineers; (iii)~actor-critic policies produce high-frequency command jitter
that degrades tire wear; and (iv)~without an explicit spatial
representation, the network must re-discover track geometry from reward
signals alone~\cite{remonda2024,kabzan2019gpmpc}.

We resolve these tensions through architectural separation: the RL policy is
restricted to the tactical layer, leaving model-based solvers for planning
and control.  A Truncated Quantile Critics (TQC) policy outputs a
low-dimensional action vector that reshapes the feasible lateral corridor
$\mathcal{C}_t=\{(s_k,n_k^-,n_k^+)\}$ and the FSM hysteresis thresholds,
without ever modifying the planner cost or the controller dynamics.
Critically, the TQC observation vector embeds a \emph{fixed-track prior}: a
frozen, map-level encoding of raceline curvature, track banking, and
friction-limit speed at each arc-length station.  This spatial prior supplies
the policy with geometric context, so the learned tactical decision is
conditioned on upcoming curvature and available track width rather than on
reward history alone.  A heuristic FSM translates the continuous TQC output
into a discrete mode sequence and a corridor-carving decision, delivering a
guidance tuple $(m_t,V_{{\rm cap},t},\mathcal{C}_t,\pi_t)$ to the
planner.

Given the corridor, the planning layer solves a sparse second-order cone
program (SOCP) for a curvature-smooth lateral path, followed by a
forward--backward speed sweep on a super-elliptic GGV envelope with
empirically identified exponent $p=0.75$.  This two-stage decomposition is
substantially faster than monolithic nonlinear trajectory
optimization~\cite{verschueren2014} while naturally capturing the coupled
longitudinal--lateral grip limits.  The resulting reference is tracked by a
nonlinear Model Predictive Contouring Controller (MPCC) that augments an
$8$-state dynamic bicycle model with a path-progress state; a terminal
progress reward automatically generates corner entry-braking and
exit-acceleration without per-corner reference shaping, and a soft rear-axle
friction-cone constraint prevents tire saturation without rendering the
nonlinear program infeasible.  All three layers share a common Frenet
representation.

The contributions of this paper are:
\begin{enumerate}
\item \textbf{Map-conditioned tactical corridor synthesis.}
A TQC policy receives a frozen preview of curvature, track width, and
friction-limited speed and outputs only interpretable tactical quantities:
side preference, safety margin, velocity cap, and corridor bounds.  A
hysteretic FSM converts this output into a time-varying feasible set
$\mathcal{C}_t$, preventing the learned policy from directly commanding
actuators or altering downstream costs.

\item \textbf{Convex local path planning with GGV-constrained speed profiling.}
A sparse SOCP over Frenet state and curvature slack produces a
curvature-regularized path inside the tactical corridor.  A
forward--backward sweep on a super-elliptic GGV envelope then constructs the
corresponding grip-limited speed profile, separating geometric feasibility
from speed feasibility while preserving their coupling through curvature.

\item \textbf{Closed-loop validation of the hierarchy in head-to-head racing.}
An $8$-state dynamic-bicycle MPCC with a terminal progress reward and a soft
rear-axle friction constraint executes the planned references.  Two
closed-loop Yas Marina fast-closure overtake simulations quantify pass time,
tactical gain, peak speed, and tracking error, showing that the same planner
and controller can execute distinct tactical corridors without case-specific
retuning.
\end{enumerate}

The remainder of the paper is organized as follows.
Section~\ref{sec:model} derives the unified vehicle and Frenet models.
Section~\ref{sec:tactical} presents the RL-driven tactical layer.
Section~\ref{sec:planner} details the SOCP path planner and the FB speed
profiler.
Section~\ref{sec:controller} formulates the MPCC controller.
Section~\ref{sec:experiments} reports two closed-loop overtake simulations.
Section~\ref{sec:conclusion} concludes.

% =====================================================================
\section{Vehicle, Track and GGV Modeling}\label{sec:model}
% =====================================================================
All subsequent layers use the same two descriptions of the vehicle state.  The
controller uses a dynamic bicycle model in an inertial frame, because tire
slip, yaw dynamics and actuator limits must be enforced at the control rate.
The tactical and planning layers use the Frenet coordinates of the same
vehicle with respect to the track centerline, because overtaking decisions are
geometric statements about longitudinal progress and lateral free space.  This
section fixes the notation used by the remainder of the paper.

\subsection{Dynamic Bicycle Model}
Let the continuous-time controller state and input be
\begin{equation}
\bm{x}=\bigl[X,\;Y,\;\psi,\;v_x,\;v_y,\;r,\;\delta,\;s_p\bigr]^\top,\qquad
\bm{u}=\bigl[a,\;\dot\delta\bigr]^\top ,
\label{eq:state}
\end{equation}
where $(X,Y)$ is the inertial position, $\psi$ is yaw angle, $(v_x,v_y)$ are
body-frame longitudinal and lateral velocities, $r$ is yaw rate, $\delta$ is
front steering angle, and $s_p$ is an auxiliary progress state used only by
the MPCC terminal reward.  The input $a$ is the commanded longitudinal
acceleration and $\dot\delta$ is the steering-rate command.  The rigid-body
kinematics are
\begin{equation}
\dot X=v_x\cos\psi-v_y\sin\psi,\quad
\dot Y=v_x\sin\psi+v_y\cos\psi,\quad
\dot\psi=r,\quad \dot s_p=v_x .
\label{eq:kin}
\end{equation}
The front and rear slip angles are
\begin{align}
\alpha_f &= \arctan\!\left(\frac{v_y+l_f r}{v_x+\varepsilon_v}\right)-\delta,
\label{eq:alphaf}\\
\alpha_r &= \arctan\!\left(\frac{v_y-l_r r}{v_x+\varepsilon_v}\right),
\label{eq:alphar}
\end{align}
where the small positive $\varepsilon_v$ regularizes the low-speed
denominator.  With linear lateral tire stiffness,
\begin{equation}
F_{yf}=-C_f\alpha_f,\qquad F_{yr}=-C_r\alpha_r .
\label{eq:tire}
\end{equation}
The resulting Newton--Euler equations are
\begin{align}
\dot v_x &= a+r v_y-\frac{1}{m}F_{yf}\sin\delta,
\label{eq:vxdot}\\
\dot v_y &= \frac{1}{m}\bigl(F_{yf}\cos\delta+F_{yr}\bigr)-r v_x,
\label{eq:vydot}\\
\dot r &= \frac{1}{I_z}\bigl(l_fF_{yf}\cos\delta-l_rF_{yr}\bigr),
\label{eq:rdot}\\
\dot\delta &= \dot\delta .
\label{eq:deltadot}
\end{align}
Equations~(\ref{eq:kin})--(\ref{eq:deltadot}) define
\begin{equation}
\dot{\bm{x}}=f_c(\bm{x},\bm{u}),\qquad
\bm{x}_{k+1}=f_d(\bm{x}_k,\bm{u}_k),
\label{eq:fd}
\end{equation}
where $f_d$ is the fourth-order Runge--Kutta discretization with sampling
time $T_s$.  Numerical parameter values are collected in
Table~\ref{tab:setup_params}.

\subsection{Track Frenet Coordinates}
Let $\bm{r}_c(s)\in\R^2$ be the track centerline parameterized by arc length
$s\in[0,L_{\rm trk})$.  Its tangent angle is $\theta(s)$ and its curvature is
$\kappa(s)=d\theta/ds$.  Define the tangent and normal unit vectors as
\begin{equation}
\bm{t}(s)=\bigl[\cos\theta(s),\;\sin\theta(s)\bigr]^\top,\qquad
\bm{n}_c(s)=\bigl[-\sin\theta(s),\;\cos\theta(s)\bigr]^\top .
\label{eq:frenet_frame}
\end{equation}
Any vehicle pose sufficiently close to the track is represented by
\begin{equation}
\bm{p}=\begin{bmatrix}X\\Y\end{bmatrix}
=\bm{r}_c(s)+n\,\bm{n}_c(s),\qquad
\xi=\psi-\theta(s),
\label{eq:frenet_pose}
\end{equation}
where $n$ is the signed lateral offset and $\xi$ is the heading error.  The
body velocity projected onto the Frenet axes is
\begin{equation}
v_s=v_x\cos\xi-v_y\sin\xi,\qquad
v_n=v_x\sin\xi+v_y\cos\xi .
\label{eq:frenet_velocity}
\end{equation}
Differentiating~(\ref{eq:frenet_pose}) gives the exact kinematic relation
\begin{align}
\dot s &= \frac{v_s}{1-\kappa(s)n},
\label{eq:sdot_frenet}\\
\dot n &= v_n,
\label{eq:ndot_frenet}\\
\dot\xi &= r-\kappa(s)\dot s .
\label{eq:xidot_frenet}
\end{align}
The wrapped progress difference used in tactical reasoning is always computed
on this same coordinate,
\begin{equation}
\Delta s=\bigl((s_o-s_e+L_{\rm trk}/2)\bmod L_{\rm trk}\bigr)-L_{\rm trk}/2,
\qquad \Delta n=n_o-n_e .
\label{eq:rel}
\end{equation}

For the path planner it is more convenient to use $s$ as the independent
variable and to describe a path by $\bm{z}(s)=[n(s),\xi(s)]^\top$.  Let
$\omega(s)$ denote the curvature-like steering variable of this geometric
path.  The spatial Frenet model is
\begin{equation}
\bm{z}'(s)=
\bm{f}_s\bigl(\bm{z}(s),\omega(s);s\bigr)
=
\begin{bmatrix}
(1-\kappa n)\tan\xi\\[2pt]
(1-\kappa n)\omega/\cos\xi-\kappa
\end{bmatrix}.
\label{eq:spatial_model}
\end{equation}
At planning iteration $j$, this model is linearized about the previous path
iterate $(\bar{\bm{z}}_k,\bar\omega_k)$ at each station $s_k$,
\begin{equation}
\bm{z}'(s_k)\approx
\bm{A}_k\bm{z}_k+\bm{B}_k\omega_k+\bm{b}_k,
\label{eq:linear_spatial_model}
\end{equation}
where $\bm{A}_k$, $\bm{B}_k$ and $\bm{b}_k$ are the Jacobian and affine
residual of $\bm{f}_s$.  This single linearized Frenet model is the one used
by the SOCP in Sec.~\ref{sec:planner}; no new lateral or heading variables are
introduced later.

\subsection{GGV Handling Envelope}
The speed planner and MPCC share the same handling-limit interpretation.  For
a speed $V$, the identified GGV table returns the available tangential and
normal acceleration limits $a_T^{\max}(V)$ and $a_N^{\max}(V)$.  Combined slip
is approximated by the super-elliptic envelope
\begin{equation}
\left|\frac{a_N}{a_N^{\max}(V)}\right|^p+
\left|\frac{a_T}{a_T^{\max}(V)}\right|^p\le1,
\label{eq:ggv}
\end{equation}
with exponent $p=0.75$.  Given a path curvature $\kappa_p$ and speed $V$, the
normal acceleration is $a_N=\kappa_pV^2$, and the remaining tangential
acceleration is
\begin{equation}
a_T^{\rm avail}(V,a_N)=a_T^{\max}(V)
\left[1-\left(\frac{|a_N|}{a_N^{\max}(V)}\right)^p\right]^{1/p}.
\label{eq:ggv_tangent}
\end{equation}
Aerodynamic drag is modeled as
\begin{equation}
a_{\rm drag}(V)=\frac{1}{m}\bigl(c_{d,2}V^2+c_{d,1}V+c_{d,0}\bigr),
\label{eq:drag}
\end{equation}
and is added with the acceleration or braking sign convention stated in
Sec.~\ref{sec:planner}.  The envelope~(\ref{eq:ggv}) is not a reward term; it
is a physical admissibility model used by both the speed profiler and the
controller.

% =====================================================================
\section{Tactical Layer: Map-Conditioned Corridor Synthesis}\label{sec:tactical}
% =====================================================================
The tactical layer is a learned decision module with a deterministic
optimization interface.  Let $\mathcal{B}(s)=[n_{\rm trk}^-(s),n_{\rm trk}^+(s)]$
denote the track boundary in the Frenet coordinates of Sec.~\ref{sec:model},
and let $\mathcal{O}_t$ denote the observed opponent set.  Online operation is
written as the composition
\begin{equation}
\begin{aligned}
\widetilde{\bm{o}}_t
&=\mathcal{E}\bigl(\bm{x}_{e,t},\mathcal{O}_t,\mathcal{P},
m_t,\sigma_t^{\rm lock}\bigr),\\
\bm{u}_t&\sim\pi_\theta(\cdot|\widetilde{\bm{o}}_t),\\
\bm{a}_t&=\Pi_{\mathcal{A}}\!\left(
\underline{\bm{a}}+\frac{\bm{u}_t+\bm{1}}{2}\odot
(\overline{\bm{a}}-\underline{\bm{a}})\right),\\
m_{t+1}&=\mathcal{F}_{m}(\bm{o}_t,\bm{a}_t,m_t,\sigma_t^{\rm lock}),\\
\mathcal{C}_t&=\mathcal{F}_{c}(\mathcal{B},\mathcal{O}_t,\bm{a}_t,
m_{t+1},\sigma_t^{\rm lock}),\\
\mathcal{G}_t&=(m_{t+1},V_{{\rm cap},t},\mathcal{C}_t,\pi_t).
\end{aligned}
\label{eq:tactical_stack}
\end{equation}
Only $\pi_\theta$ is learned.  The projection $\Pi_{\mathcal A}$, the
finite-state machine $\mathcal{F}_m$, and the corridor carver
$\mathcal{F}_c$ are deterministic maps that convert the policy output into
auditable constraints for the planner.  The training problem is
\begin{equation}
\pi_\theta^\star=
\operatorname*{arg\,max}_{\pi_\theta}
\mathbb{E}_{\pi_\theta}\!
\left[
\sum_{t=0}^{T_{\rm ep}-1}\gamma^t
r\bigl(\widetilde{\bm{o}}_t,\bm{a}_t,m_t,\mathcal{C}_t\bigr)
\right],
\label{eq:tactical_objective}
\end{equation}
so reinforcement learning is used to select a tactical feasible set rather
than low-level steering or throttle.

The encoder $\mathcal{E}$ combines a duel-state observation, a fixed track
prior, and a vehicle-capability prior.  The dynamic observation is
\begin{align}
\bm{o}_t^{\rm ego}
&=\bigl[
\bar V_e,\;\bar\xi_e,\;\bar n_e,\;\bar a_{T,e},\;\bar a_{N,e},\;\bar s_e
\bigr]^\top,
\label{eq:obs_ego}\\
\bm{o}_t^{\rm opp}
&=\bigl[
\bar{\Delta s},\;\bar{\Delta n},\;\bar{\Delta V_s},\;\bar s_o,\;\bar n_o,\;
\bar g,\;\bar\tau_{\rm tc},\;\bar V_o
\bigr]^\top ,
\label{eq:obs_opp}\\
\bm{o}_t
&=\bigl[
\bm{o}_t^{\rm ego},\;\bm{o}_t^{\rm opp},\;\bm{b}_t,\;
\sigma_t^{\rm lock},\;\bm{e}_{m_t}
\bigr].
\label{eq:obs}
\end{align}
With five FSM modes, $\bm{o}_t\in\mathbb{R}^{23}$.  Each scalar is normalized
as
\begin{equation}
\bar z=\operatorname{sat}(z/z_{\rm char}),\qquad
\operatorname{sat}(q)=\min\{1,\max\{-1,q\}\},
\label{eq:normalization}
\end{equation}
using the characteristic values in Table~\ref{tab:setup_params}.  The
relative progress and lateral separation are given by~(\ref{eq:rel}),
$\Delta V_s=V_o-V_e$, and $g=\sqrt{\Delta s^2+\Delta n^2}$.  The clipped
time-to-collision feature is
\begin{equation}
\tau_{\rm tc}=
\begin{cases}
\operatorname{clip}\!\left(\Delta s/\max(-\Delta V_s,\varepsilon_\tau),
0,\tau_{\rm tc,\max}\right), & \Delta s>0,\ \Delta V_s<0,\\
\tau_{\rm tc,\max}, & \text{otherwise},
\end{cases}
\label{eq:ttc}
\end{equation}
so only a closing leader creates a finite urgency scale.  The bucket vector
$\bm{b}_t$ coarsely encodes whether the ego is fast-closing, speed-matched or
receding.  The memory variables $\sigma_t^{\rm lock}$ and $\bm{e}_{m_t}$ make
side commitment and hysteresis observable to the policy.

The dynamic duel state is augmented by a non-trainable map
$\mathcal{P}=\{\bm{\rho}_i\}_{i=0}^{M-1}$.  This map contains not only
curvature and width, but also a local estimate of vehicle capability.  From
the GGV envelope in Sec.~\ref{sec:model}, define the curvature-limited speed
preview
\begin{equation}
V_{\rm ggv}(s)=
\sup_{0\le V\le v_{x,\max}}
\left\{
V:\; |\kappa(s)|V^2\le a_N^{\max}(V)
\right\},
\label{eq:ggv_preview}
\end{equation}
with $V_{\rm ggv}=v_{x,\max}$ when the curvature constraint is inactive.
For a preview interval
$\mathcal I(s;d_a,d_b)=[s+d_a,s+d_b]$, the statistics inserted into the
policy are
\begin{equation}
\kappa_{d_a:d_b}^{\max}(s)=
\max_{\ell\in\mathcal I(s;d_a,d_b)}|\kappa(\ell)|,\quad
W_{\min}^{L/R}(s)=
\min_{\ell\in[s,s+d_p]} |n_{\rm trk}^{+/-}(\ell)|,\quad
V_{\min}^{\rm ggv}(s)=
\min_{\ell\in[s,s+d_p]}V_{\rm ggv}(\ell).
\label{eq:prior_statistics}
\end{equation}
Querying $\mathcal{P}$ at progress $s$ returns
\begin{equation}
\begin{aligned}
\bm{\rho}(s)=
\bigl[&
\bar\kappa_0,\;
\bar\kappa_{0:d_1}^{\max},\;
\bar\kappa_{d_1:d_2}^{\max},\;
\bar\kappa_{d_2:d_3}^{\max},\;
\bar\kappa_{d_3:d_p}^{\max},\;
\bar\kappa_{\rm mean},\;
\bar\kappa_{\rm std},\\
&\bar W_{\min}^{L},\;
\bar W_{\min}^{R},\;
\bar W_0^L,\;
\bar W_0^R,\;
\bar{\Delta W}_{\min},\;
\bar V_0^{\rm ggv},\;
\bar V_{\min}^{\rm ggv}
\bigr]^\top .
\end{aligned}
\label{eq:track_prior}
\end{equation}
The curvature terms encode how rapidly the track is about to turn, the width
terms encode whether an overtake side remains open, and the GGV terms encode
whether that side can be exploited by the present vehicle without exceeding
normal-acceleration capability.  The final actor--critic input is
\begin{equation}
\widetilde{\bm{o}}_t=
\bigl[\bm{o}_t,\bm{\rho}(s_e),\bm{\rho}(s_o)\bigr]\in\mathbb{R}^{51},
\label{eq:track_aug_obs}
\end{equation}
so an identical relative gap can yield different actions on a straight, at
corner entry, or before a narrow exit.

The actor output is not a trajectory but a ten-dimensional tactical parameter
vector.  A sample $\bm{u}_t\in[-1,1]^{10}$ is mapped to
\begin{equation}
\bm{a}_t=\Pi_{\mathcal A}\!\left(
\underline{\bm{a}}+
\frac{1}{2}(\bm{u}_t+\bm{1})\odot(\overline{\bm{a}}-\underline{\bm{a}})
\right),
\label{eq:tanh_map}
\end{equation}
\begin{equation}
\bm{a}_t=
\bigl[
L_t^{\rm cap},R_t^{\rm cap},d_{\rm safe},V_{\rm cap},
g_{\rm chase},g_{\rm ot},g_{\rm abort},b_{\rm lat},w_{\rm safe},q_{\rm side}
\bigr]^\top .
\label{eq:action_vec}
\end{equation}
The projection $\Pi_{\mathcal A}$ enforces
\begin{equation}
\mathcal A=\left\{
\bm{a}:\;
g_{\rm chase}\ge g_{\rm ot}+\varepsilon_{co},\;
g_{\rm abort}\ge g_{\rm chase}+\varepsilon_{ca},\;
L_t^{\rm cap}+R_t^{\rm cap}\ge\varepsilon_{\rm corr}
\right\}.
\label{eq:safety_clip}
\end{equation}
The components of $\bm{a}_t$ have direct downstream meanings:
$(L_t^{\rm cap},R_t^{\rm cap})$ cap the left/right corridor width,
$d_{\rm safe}$ sets opponent clearance, $V_{\rm cap}$ caps the speed profile,
$(g_{\rm chase},g_{\rm ot},g_{\rm abort})$ tune FSM hysteresis, $b_{\rm lat}$
sets the far-field lateral bias, $w_{\rm safe}$ controls recovery width, and
$q_{\rm side}$ chooses the pass side,
\begin{equation}
\sigma_t=
\begin{cases}
+1, & q_{\rm side}\ge0\quad \text{left-side pass},\\
-1, & q_{\rm side}<0\quad \text{right-side pass}.
\end{cases}
\label{eq:side_sign}
\end{equation}
Once $\sigma_t$ is accepted in \textsc{Overtake}, it is copied into
$\sigma_t^{\rm lock}$ and held until the duel terminates.

The continuous tactical action is converted into a persistent intention by a
hysteretic finite-state machine with modes
$\mathcal{M}=\{\textsc{Raceline},\textsc{Shadow},\textsc{Overtake},
\textsc{Hold},\textsc{Defend}\}$.  Let $g_{\rm rear}$ denote the closest
rear gap and $\bar c_t$ indicate that the post-overtake immunity window has
expired.  The selected side is considered ready when the previewed side width
and GGV speed are both admissible,
\begin{equation}
{\rm ready}_t(\sigma)=
\mathbb{1}\!\left[
\min_{s\in[s_e,s_e+d_p]} W_\sigma(s) \ge L_{\min}^{\rm corr}
\right]
\mathbb{1}\!\left[
\min_{s\in[s_e,s_e+d_p]} V_{\rm ggv}(s) \ge \min\{V_e,V_{\rm cap}\}
\right],
\label{eq:ready}
\end{equation}
where $W_{+1}$ and $W_{-1}$ are the left/right widths remaining after
opponent exclusion.  The mode update is
\begin{equation}
m_{t+1}=
\begin{cases}
\textsc{Overtake}, &
g\le g_{\rm ot}\wedge {\rm ready}_t(\sigma_t)\wedge\bar c_t,\\
\textsc{Shadow}, &
g_{\rm ot}<g\le g_{\rm chase},\\
\textsc{Hold}, &
m_t=\textsc{Overtake}\wedge{\rm abort}_t,\\
\textsc{Defend}, &
g>g_{\rm def}^{\rm front}\wedge0<g_{\rm rear}<g_{\rm def}^{\rm rear}\wedge\bar c_t,\\
\textsc{Raceline}, &
g>g_{\rm chase},\\
m_t, & \text{otherwise}.
\end{cases}
\label{eq:fsm}
\end{equation}
The abort flag is
\begin{equation}
{\rm abort}_t=
\bigl(g>g_{\rm abort}\bigr)\vee
\bigl(\Delta n_{\min}<\Delta n_{\rm safe}\wedge |\Delta s|<g_{\rm interlock}\bigr),
\label{eq:abort}
\end{equation}
where $\Delta n_{\min}$ is the minimum lateral separation predicted within
the local interaction window.  Therefore the policy adjusts thresholds and
side preference, while the hybrid logic preserves a persistent mode sequence
for the downstream optimizer.

The reward is defined on tactical outcome and constraint quality.  At each
step,
\begin{equation}
\begin{aligned}
r_t={}&
w_{\rm ot}\phi_{\rm ot}
+w_{\rm ovb}\phi_{\rm ovb}
+w_{\rm col}\phi_{\rm col}
+w_{\rm lat}\phi_{\rm lat}
+w_{\rm crush}\phi_{\rm crush}
+w_{\rm app}\phi_{\rm app}\\
&+
w_{\rm sm}\phi_{\rm sm}
+w_{\rm gs}\phi_{\rm gs}
+w_{\rm sbs}\phi_{\rm sbs}
+w_{\rm wide}\phi_{\rm wide},
\end{aligned}
\label{eq:reward_sum}
\end{equation}
with weights listed in Table~\ref{tab:setup_params}.  The event and geometry
features are
\begin{align}
\phi_{\rm ot}
&=\mathbb{1}[\Delta s_{t-1}>0]\mathbb{1}[\Delta s_t<0]\mathbb{1}[\bar c_t],
\label{eq:reward_pass}\\
\phi_{\rm ovb}
&=\mathbb{1}[\Delta s_{t-1}<0]\mathbb{1}[\Delta s_t>0],
\label{eq:reward_overtaken}\\
\phi_{\rm col}
&=\mathbb{1}[{\rm contact}\vee{\rm offtrack}\vee{\rm comm\_loss}],
\label{eq:reward_collision}\\
\phi_{\rm lat}
&=\mathbb{1}[m_t=\textsc{Raceline}]
\left(\frac{n_e-n_{\rm rl}(s_e)}{n_{\rm char}}\right)^2,
\label{eq:reward_lateral}\\
\phi_{\rm crush}
&=\mathbb{1}[|\Delta s|<g_{\rm interlock}]
\left(\frac{\max\{0,\Delta n_{\rm safe}-|\Delta n|\}}{n_{\rm char}}\right)^2 .
\label{eq:reward_crush}
\end{align}
The approach feature is
\begin{equation}
\phi_{\rm app}=
\mathbb{1}[\Delta s>0]\,
\mathbb{1}[g_t<g_{\rm chase}]\,
\operatorname{clip}\!\left(\frac{g_{t-1}-g_t}{g_{\rm char}},0,1\right),
\label{eq:approach_reward}
\end{equation}
and the regularizers applied to the policy output and the published corridor
are
\begin{equation}
\phi_{\rm sm}=
\|\bm{a}_t-\bm{a}_{t-1}\|_{\bm{W}_a}^2,
\label{eq:action_smooth_reward}
\end{equation}
\begin{equation}
\phi_{\rm gs}=
\frac{1}{N_{\rm sta}}\sum_{k=0}^{N_{\rm sta}-1}
\Bigl[
(n_{k,t}^+-n_{k,t-1}^+)^2+
(n_{k,t}^--n_{k,t-1}^-)^2
\Bigr],
\label{eq:corridor_smooth_reward}
\end{equation}
\begin{align}
\phi_{\rm sbs}
&=\mathbb{1}[|\Delta s|<g_{\rm interlock}]
\mathbb{1}[|\Delta n|\ge\Delta n_{\rm safe}],
\label{eq:reward_sbs}\\
\phi_{\rm wide}
&=\frac{1}{N_{\rm sta}}\sum_{k=0}^{N_{\rm sta}-1}
\operatorname{clip}\!\left(
\frac{n_k^+-n_k^- -L_{\min}^{\rm corr}}{n_{\rm char}},0,1
\right).
\label{eq:reward_wide}
\end{align}
Positive weights reward pass completion, useful gap closing, safe
side-by-side tenure and non-degenerate corridors.  Negative weights penalize
being overtaken, collision, unnecessary raceline deviation, lateral
compression, action jitter and corridor jitter.  Hence the reward shapes only
the tactical interface; it does not modify the SOCP or MPCC objectives.

The policy is optimized with Truncated Quantile Critics.  Two critics
parameterized by $\varphi_i$ approximate the soft return distribution with
atoms $\{z_{\varphi_i}^{j}(\widetilde{\bm{o}},\bm{u})\}_{j=1}^{N_q}$.  Given
a replay transition, draw
$\bm{u}'\sim\pi_\theta(\cdot|\widetilde{\bm{o}}')$ and form
\begin{equation}
\mathcal{Z}_{\rm targ}=
\operatorname{sort}\Bigl(
\{r_t+\gamma(1-d_t)
[z_{\bar\varphi_i}^{j}(\widetilde{\bm{o}}',\bm{u}')
-\alpha\log\pi_\theta(\bm{u}'|\widetilde{\bm{o}}')]\}_{i,j}
\Bigr)_{1:K}.
\label{eq:tqc_target}
\end{equation}
Retaining the lowest $K$ atoms reduces overestimation of rare, unsafe passes.
The critic minimizes
\begin{equation}
J_Q(\varphi)=
\sum_{i=1}^{2}\sum_{j=1}^{N_q}
\mathbb{E}_{\mathcal D}
\left[
\rho_{\hat\tau_j}^{\kappa_q}
\left(
y-z_{\varphi_i}^{j}(\widetilde{\bm{o}}_t,\bm{u}_t)
\right)
\right],
\label{eq:tqc_critic}
\end{equation}
where $y\in\mathcal{Z}_{\rm targ}$, and the actor minimizes
\begin{equation}
J_\pi(\theta)=
\mathbb{E}_{\widetilde{\bm{o}}\sim\mathcal D,\bm{u}\sim\pi_\theta}
\left[
\alpha\log\pi_\theta(\bm{u}|\widetilde{\bm{o}})
-\frac{1}{K}\sum_{z\in\mathcal{Z}_{K}(\widetilde{\bm{o}},\bm{u})}z
\right],
\label{eq:tqc_actor}
\end{equation}
where $\mathcal{Z}_{K}$ denotes the truncated critic atoms at
$(\widetilde{\bm{o}},\bm{u})$.  The entropy temperature $\alpha$ is tuned
automatically.  Training episodes randomize initial gap, relative speed,
lateral offset and opponent behavior; termination occurs at pass completion,
being overtaken, collision, off-track excursion, communication loss or
timeout.  Evaluation uses the deterministic mean action with the same FSM and
corridor carver.

The tactical output consumed by the planner is
\begin{equation}
\mathcal{G}_t=\bigl(m_t,V_{{\rm cap},t},\mathcal{C}_t,\pi_t\bigr),\qquad
\mathcal{C}_t=\{(s_k,n_k^-,n_k^+)\}_{k=0}^{N_{\rm sta}-1},
\label{eq:guidance_tuple}
\end{equation}
where $\pi_t$ stores race-control permission.  The carver is the deterministic
map
\begin{equation}
\mathcal{C}_t=
\mathcal{F}_{c}
\left(
\{n_{\rm trk}^-(s_k),n_{\rm trk}^+(s_k)\}_{k=0}^{N_{\rm sta}-1},
\mathcal{O}_t,m_t,\sigma_t^{\rm lock},\bm{a}_t
\right).
\label{eq:carver_map}
\end{equation}
It first initializes the station grid and physical bounds,
\begin{equation}
s_k=s_e+k\Delta s_{\rm sta},\qquad
n_k^-=n_{\rm trk}^-(s_k),\qquad
n_k^+=n_{\rm trk}^+(s_k).
\label{eq:corridor_init}
\end{equation}
For an opponent with heading error $\xi_o$ and footprint $(L_v,W_v)$, the
half-extents projected onto Frenet axes are
\begin{equation}
\rho_n=\frac{1}{2}(L_v|\sin\xi_o|+W_v|\cos\xi_o|),\qquad
\rho_s=\frac{1}{2}(L_v|\cos\xi_o|+W_v|\sin\xi_o|),
\label{eq:footprint_projection}
\end{equation}
and the lateral exclusion radius is
\begin{equation}
\rho_{\rm exc}=\rho_n+\frac{W_v}{2}+d_{\rm safe}.
\label{eq:rho_exc}
\end{equation}
The longitudinal influence of the footprint is
\begin{equation}
\phi_{\rm fade}(\Delta s_k)=
\begin{cases}
\cos^{p_{\rm fade}}\!\left(\frac{\pi}{2}\frac{|\Delta s_k|}{S_{\rm fade}}\right),
& |\Delta s_k|<S_{\rm fade},\\
0, & \text{otherwise},
\end{cases}
\label{eq:fade}
\end{equation}
multiplied by a startup ramp $\phi_{\rm sr}(k)$ and a distance ramp
$\phi_{\rm rmp}(k)$.  In \textsc{Overtake}, the bound on the opponent side is
tightened and the selected side is left open:
\begin{equation}
\begin{cases}
n_k^- \leftarrow
\max\!\left(n_k^-,\, n_{o,k}+\rho_{\rm exc}\phi_{\rm fade}\phi_{\rm sr}\phi_{\rm rmp}\right),
& \sigma_t=+1,\\[3pt]
n_k^+ \leftarrow
\min\!\left(n_k^+,\, n_{o,k}-\rho_{\rm exc}\phi_{\rm fade}\phi_{\rm sr}\phi_{\rm rmp}\right),
& \sigma_t=-1.
\end{cases}
\label{eq:overtake_bound}
\end{equation}
In \textsc{Shadow}, the interval is shaped into an asymmetric funnel toward
$\sigma_t$; in \textsc{Hold}, it is relaxed toward the raceline; in
\textsc{Defend}, it is biased toward the defensive side.  The learned bias
and width caps are then blended into the far field.  With
\begin{equation}
w_k=
\begin{cases}
(k/k_{\rm blend})^2, & k<k_{\rm blend},\\
1, & k\ge k_{\rm blend},
\end{cases}
\label{eq:rl_blend_w}
\end{equation}
\begin{align}
n_k^+ &\leftarrow (1-w_k)n_k^+
+w_k\min\bigl(n_k^+,b_{\rm lat}+L_t^{\rm cap}\bigr),
\label{eq:rl_blend_max}\\
n_k^- &\leftarrow (1-w_k)n_k^-
+w_k\max\bigl(n_k^-,b_{\rm lat}-R_t^{\rm cap}\bigr).
\label{eq:rl_blend_min}
\end{align}
The minimum-width guard is
\begin{equation}
n_k^+-n_k^- \ge L_{\min}^{\rm corr}.
\label{eq:min_corridor_width}
\end{equation}
If this inequality cannot be recovered within the physical track boundary,
the previous feasible interval is held at that station.  Finally, the
published corridor is temporally filtered,
\begin{equation}
\begin{bmatrix}n_k^-\\n_k^+\end{bmatrix}_{t}
\leftarrow
\beta
\begin{bmatrix}n_k^-\\n_k^+\end{bmatrix}_{t}
+(1-\beta)
\begin{bmatrix}n_k^-\\n_k^+\end{bmatrix}_{t-1},
\label{eq:ema}
\end{equation}
and the tuple $\mathcal{G}_t$ is sent to the SOCP planner.  Thus the learned
layer contributes exactly three planning quantities: $m_t$ selects the
mode-dependent terminal target, $V_{{\rm cap},t}$ clips the GGV speed profile,
and $\mathcal{C}_t$ supplies the lateral inequalities
$n_k^-\le n_k\le n_k^+$ in Sec.~\ref{sec:planner}.  If a tactical update is
delayed by the semi-physical interface, the last valid $\mathcal{G}_t$ is
held for one planning cycle.

% =====================================================================
\section{Local Planning}\label{sec:planner}
% =====================================================================
The local planner converts the tactical tuple
$\mathcal{G}_t=(m_t,V_{{\rm cap},t},\mathcal{C}_t,\pi_t)$ into a
time-parameterized reference.  Its structure is deliberately split:
\begin{equation}
\mathcal{G}_t
\xrightarrow{\;\text{ECOS SOCP}\;}
\{s_k,n_k^\star,\xi_k^\star,\kappa_{p,k}^\star\}_{k=0}^{N-1}
\xrightarrow{\;\text{FB speed surface}\;}
\mathcal{R}_t .
\label{eq:planner_pipeline}
\end{equation}
The first stage solves a sparse second-order cone program for the Frenet path.
The second stage computes a forward--backward (FB) speed surface on the
resulting local path using the GGV envelope of Sec.~\ref{sec:model}.

The horizon is sampled at
\begin{equation}
s_k=s_e+k\Delta s_{\rm sta},\qquad k=0,\ldots,N-1,
\label{eq:planner_grid}
\end{equation}
where the wrapped value of $s_k$ is used when querying track data.  The
tactical corridor is intersected with the physical track boundary,
\begin{equation}
\underline n_k=\max\{n_{\rm trk}^-(s_k),n_k^-\},\qquad
\overline n_k=\min\{n_{\rm trk}^+(s_k),n_k^+\}.
\label{eq:planner_corridor_intersection}
\end{equation}
If no tactical corridor is active, $\underline n_k$ and $\overline n_k$ are
the track limits after the standard safety margin.  The optimization variables
are
\begin{equation}
\bm y=
\left[
\{n_k,\xi_k,\omega_k,\sigma_k^\omega\}_{k=0}^{N-1},
\epsilon_n,\epsilon_\xi
\right],
\qquad
\bm z_k=\begin{bmatrix}n_k\\ \xi_k\end{bmatrix},
\label{eq:path_decision}
\end{equation}
where $(n_k,\xi_k)$ are the Frenet variables of Sec.~\ref{sec:model},
$\omega_k$ is the curvature-like path input in~(\ref{eq:spatial_model}),
$\sigma_k^\omega$ is the cone epigraph for $\omega_k^2$, and
$(\epsilon_n,\epsilon_\xi)$ are terminal relaxation variables.

The initial condition is hard,
\begin{equation}
\bm z_0=\begin{bmatrix}n_e\\ \xi_e\end{bmatrix},
\label{eq:path_initial}
\end{equation}
while the terminal target is mode dependent,
\begin{equation}
n_T=
\begin{cases}
n_{\rm rl}(s_{N-1}), &
m_t=\textsc{Raceline},\;
n_{\rm rl}(s_{N-1})\in[\underline n_{N-1},\overline n_{N-1}],\\[2pt]
\frac12(\underline n_{N-1}+\overline n_{N-1}), & \text{otherwise},
\end{cases}
\qquad \xi_T=0 .
\label{eq:terminal_mode}
\end{equation}
Instead of forcing this terminal condition as an equality, the implemented
planner uses soft terminal inequalities,
\begin{equation}
\begin{aligned}
-\epsilon_n\le n_{N-1}-n_T\le \epsilon_n,\qquad
-\epsilon_\xi\le \xi_{N-1}-\xi_T\le \epsilon_\xi,\\
\epsilon_n\ge0,\qquad \epsilon_\xi\ge0 .
\end{aligned}
\label{eq:terminal_slack}
\end{equation}
This prevents ECOS infeasibility when a short horizon begins inside a tight
overtake corridor but still biases the path toward the tactical terminal
intent.

At sequential-linearization iteration $j$, the nonlinear spatial model
(\ref{eq:spatial_model}) is linearized at
$(\bm z_k^{j},\omega_k^{j})$,
\begin{equation}
\bm f_s(\bm z_k,\omega_k;s_k)
\approx
\bm A_k^j\bm z_k+\bm B_k^j\omega_k+\bm b_k^j
\triangleq \hat{\bm f}_{s,k}^j(\bm z_k,\omega_k).
\label{eq:path_affine_model}
\end{equation}
Trapezoidal integration gives the linear equality constraint
\begin{equation}
\bm z_{k+1}-\bm z_k
=\frac{\Delta s_k}{2}
\left[
\hat{\bm f}_{s,k}^{j}(\bm z_k,\omega_k)
+\hat{\bm f}_{s,k+1}^{j}(\bm z_{k+1},\omega_{k+1})
\right],
\quad k=0,\ldots,N-2 .
\label{eq:trap}
\end{equation}
The lateral corridor and curvature epigraph are
\begin{align}
\underline n_k\le n_k\le \overline n_k,\qquad k=0,\ldots,N-1,
\label{eq:corr_ineq}\\
\left\|
\begin{bmatrix}
2\omega_k\\ \sigma_k^\omega-1
\end{bmatrix}
\right\|_2
\le \sigma_k^\omega+1,\qquad k=0,\ldots,N-1,
\label{eq:cone}
\end{align}
which is equivalent to $(\omega_k)^2\le\sigma_k^\omega$ with
$\sigma_k^\omega\ge0$.  The SOCP solved by ECOS is
\begin{equation}
\begin{aligned}
\bm y^{j+1}=\argmin_{\bm y}\;&
\sum_{k=0}^{N-1} w_\omega \sigma_k^\omega
+w_{n,T}\epsilon_n+w_{\xi,T}\epsilon_\xi\\
\text{s.t.}\;&
\text{(\ref{eq:path_initial}), (\ref{eq:terminal_slack}),
(\ref{eq:trap}), (\ref{eq:corr_ineq}), and (\ref{eq:cone})}.
\end{aligned}
\label{eq:socp}
\end{equation}
Equivalently, ECOS receives the conic form
\begin{equation}
\min_{\bm y}\bm c^\top\bm y,\qquad
\bm A_{\rm eq}\bm y=\bm b_{\rm eq},\quad
\bm G_{\rm lp}\bm y\le\bm h_{\rm lp},\quad
\|\bm G_{q,k}\bm y+\bm g_{q,k}\|_2\le
\bm h_{q,k}^\top\bm y+h_{q,k}.
\label{eq:ecos_form}
\end{equation}
After each solve, the linearization trajectory is updated.  The loop stops
when
\begin{equation}
\Delta_n^j=\max_k|n_k^{j+1}-n_k^j|\le\varepsilon_n^{\rm sl},\qquad
\Delta_\xi^j=\max_k|\xi_k^{j+1}-\xi_k^j|\le\varepsilon_\xi^{\rm sl},
\label{eq:slin_conv}
\end{equation}
or when the iteration limit is reached.  The optimized Cartesian path and its
heading are
\begin{equation}
\bm p_k^\star=\bm r_c(s_k)+n_k^\star\bm n_c(s_k),\qquad
\psi_k^\star=\theta(s_k)+\xi_k^\star,\qquad
\kappa_{p,k}^\star=\omega_k^\star .
\label{eq:path_cart}
\end{equation}
The path arclength used by the speed planner is then recomputed from the
optimized Cartesian points,
\begin{equation}
\ell_0=0,\qquad
\ell_{k+1}=\ell_k+\|\bm p_{k+1}^\star-\bm p_k^\star\|_2,\qquad
\Delta\ell_k=\ell_{k+1}-\ell_k .
\label{eq:path_arclength}
\end{equation}

The FB speed profiler computes the largest speed surface that remains
consistent with curvature, longitudinal acceleration and braking.  First, a
curvature speed ceiling is obtained at each station:
\begin{equation}
V_{{\rm lim},k}=
\begin{cases}
v_{x,\max}, & |\kappa_{p,k}^\star|<\varepsilon_\kappa,\\[2pt]
\displaystyle
\lim_{r\to\infty}V_k^{(r)}, &
V_k^{(r+1)}=
\min\!\left\{
v_{x,\max},
\sqrt{\frac{a_N^{\max}(V_k^{(r)})}{|\kappa_{p,k}^\star|}}
\right\},
\end{cases}
\label{eq:vlim_iter}
\end{equation}
initialized with $V_k^{(0)}=v_{x,\max}$.  For any candidate speed $V$ at
station $k$, define
\begin{equation}
a_{N,k}(V)=|\kappa_{p,k}^\star|V^2,\qquad
\chi_k(V)=
\left[
1-\left(\frac{a_{N,k}(V)}{a_N^{\max}(V)}\right)^p
\right]_+^{1/p},
\label{eq:ggv_factor}
\end{equation}
where $[\cdot]_+=\max\{0,\cdot\}$ and $p$ is the GGV exponent in
(\ref{eq:ggv}).  The available forward and braking accelerations are
\begin{align}
a_{T,k}^{+}(V)
&=a_T^{+,\max}(V)\chi_k(V)+a_{\rm drag}(V),
\label{eq:fb_accel_available}\\
a_{T,k}^{-}(V)
&=a_T^{-,\max}(V)\chi_k(V)-a_{\rm drag}(V),
\label{eq:fb_brake_available}
\end{align}
where $a_T^{+,\max}$ and $a_T^{-,\max}$ are the calibrated driving and
braking envelopes.  Under the sign convention of~(\ref{eq:drag}),
$a_{\rm drag}$ is usually negative at racing speeds; hence it reduces forward
acceleration and assists braking.

The forward sweep starts from the measured speed,
\begin{equation}
V_0^f=\min\{\max(V_e,V_{\min}),V_{{\rm lim},0}\},
\label{eq:forward_init}
\end{equation}
and propagates
\begin{equation}
V_{k+1}^{f}
=
\operatorname{clip}_{[V_{\min},\,V_{{\rm lim},k+1}]}
\left(
\sqrt{(V_k^f)^2+2\Delta\ell_k\,a_{T,k}^{+}(V_k^f)}
\right),
\quad k=0,\ldots,N-2 .
\label{eq:vfwd}
\end{equation}
The backward sweep is initialized by
\begin{equation}
V_{N-1}^{b}=
\begin{cases}
\min\{V_T,V_{{\rm lim},N-1}\}, & \text{if a terminal speed is enforced},\\
\min\{V_{N-1}^{f},V_{{\rm lim},N-1}\}, & \text{otherwise},
\end{cases}
\label{eq:backward_init}
\end{equation}
and propagates
\begin{equation}
V_{k}^{b}
=
\operatorname{clip}_{[V_{\min},\,V_{{\rm lim},k}]}
\left(
\sqrt{(V_{k+1}^{b})^2+2\Delta\ell_k\,a_{T,k+1}^{-}(V_{k+1}^{b})}
\right),
\quad k=N-2,\ldots,0 .
\label{eq:vbwd}
\end{equation}
The final speed surface is the pointwise feasible envelope
\begin{equation}
V_k^\star=\min\{V_k^f,V_k^b,V_{{\rm lim},k},V_{{\rm cap},t}\}.
\label{eq:vmerge}
\end{equation}
The accelerations stored with the local reference are
\begin{equation}
a_{T,k}^\star=
\frac{(V_{k+1}^\star)^2-(V_k^\star)^2}{2\Delta\ell_k},\qquad
a_{N,k}^\star=\kappa_{p,k}^\star(V_k^\star)^2 .
\label{eq:atan}
\end{equation}
Finally, the time stamps are reconstructed by trapezoidal time integration,
\begin{equation}
\tau_0^r=0,\qquad
\tau_{k+1}^r=\tau_k^r+
\frac{2\Delta\ell_k}{V_k^\star+V_{k+1}^\star+\varepsilon_v},
\label{eq:time_reconstruct}
\end{equation}
and the planner publishes
\begin{equation}
\mathcal{R}_t=
\{(X_k^r,Y_k^r,\psi_k^r,\kappa_{p,k}^r,V_k^r,\tau_k^r)\}_{k=0}^{N_{\rm out}-1},
\label{eq:reference_tuple}
\end{equation}
after resampling the path, speed and time arrays to the controller cadence.

% =====================================================================
\section{Model Predictive Contouring Control}\label{sec:controller}
% =====================================================================
The control layer converts the local reference
$\mathcal{R}_t$ in~(\ref{eq:reference_tuple}) into the actuator command
actually applied to the race car.  The tactical layer determines the
admissible side of the overtake and the SOCP layer computes a feasible path;
the controller therefore does not change the corridor.  It solves a compact
receding-horizon Model Predictive Contouring Control (MPCC) problem in a
vehicle-fixed local frame, and then maps the first optimal speed--steering
pair into throttle, brake pressure, steering command and gear.

At control instant $t$, the current vehicle pose defines a rigid transform
$\mathcal{T}_t$ from the global frame to the local controller frame.  The
measured pose becomes the origin,
\begin{equation}
\bm{\chi}_{0|t}=\bigl[x_{0|t},y_{0|t},\theta_{0|t}\bigr]^\top
=\bm{0},
\qquad
\bm{\nu}_{-1|t}=\bigl[V_{-1|t},\delta_{-1|t}\bigr]^\top ,
\label{eq:mpcc_initial}
\end{equation}
where $V_{-1|t}$ and $\delta_{-1|t}$ are the speed and front-wheel steering
commands applied in the previous cycle.  The reference sequence used by the
controller is obtained by transforming the planner output:
\begin{equation}
\bar{\bm{r}}_{k|t}=
\bigl[x_{k|t}^{r},y_{k|t}^{r},\theta_{k|t}^{r},V_{k|t}^{r},
\delta_{k|t}^{r}\bigr]^\top ,
\qquad k=1,\ldots,N_c ,
\label{eq:mpcc_reference}
\end{equation}
with the steering feedforward reconstructed from the planned yaw-rate
reference,
\begin{equation}
r_{k|t}^{r}=\kappa_{p,k|t}^{r}V_{k|t}^{r},\qquad
\delta_{k|t}^{r}=
\arctan\!\left(\frac{L_c r_{k|t}^{r}}
{V_{k|t}^{r}+\varepsilon_v}\right).
\label{eq:steer_ff_ref}
\end{equation}
Here $L_c$ is the wheelbase used by the controller.  The local state is a
projection of the vehicle model in Sec.~\ref{sec:model} onto pose, and the
input is the commanded path speed and front steering angle:
\begin{equation}
\bm{\chi}_{k|t}=\bigl[x_{k|t},y_{k|t},\theta_{k|t}\bigr]^\top,\qquad
\bm{\nu}_{k|t}=\bigl[V_{k|t},\delta_{k|t}\bigr]^\top .
\label{eq:mpcc_state_input}
\end{equation}
The prediction model is
\begin{equation}
\bm{\chi}_{k+1|t}
=\bm{\chi}_{k|t}+T_c
\begin{bmatrix}
V_{k|t}\cos\theta_{k|t}\\
V_{k|t}\sin\theta_{k|t}\\
\Omega(V_{k|t},\delta_{k|t})
\end{bmatrix},
\label{eq:mpcc_model}
\end{equation}
where the identified yaw-rate map is
\begin{equation}
\Omega(V,\delta)=
\frac{\tan(a_\delta\delta+b_\delta)}{L_c}\,
\bigl(c_VV+d_V\bigr).
\label{eq:yaw_rate_fit}
\end{equation}
This model is deliberately low-dimensional: the GGV sweep in
Sec.~\ref{sec:planner} has already produced a speed-feasible local path, and
the controller must solve fast enough to reject pose error at the actuation
rate.

For each predicted point, define
\begin{equation}
\Delta\bm{p}_{k|t}=
\begin{bmatrix}
x_{k|t}-x_{k|t}^{r}\\
y_{k|t}-y_{k|t}^{r}
\end{bmatrix},\quad
\bm{t}_{k|t}^{r}=
\begin{bmatrix}
\cos\theta_{k|t}^{r}\\
\sin\theta_{k|t}^{r}
\end{bmatrix},\quad
\bm{n}_{k|t}^{r}=
\begin{bmatrix}
-\sin\theta_{k|t}^{r}\\
\cos\theta_{k|t}^{r}
\end{bmatrix}.
\label{eq:mpcc_frame}
\end{equation}
The contouring, lag, heading and input-tracking errors are
\begin{align}
e_{c,k|t} &=(\bm{n}_{k|t}^{r})^\top\Delta\bm{p}_{k|t},
\label{eq:mpcc_ec}\\
e_{\ell,k|t} &=(\bm{t}_{k|t}^{r})^\top\Delta\bm{p}_{k|t},
\label{eq:mpcc_el}\\
e_{\psi,k|t} &=\operatorname{wrap}
\bigl(\theta_{k|t}-\theta_{k|t}^{r}\bigr),
\label{eq:mpcc_epsi}\\
\bm{e}_{\nu,k|t} &=\bm{\nu}_{k|t}
-\bm{\nu}_{k|t}^{r},\qquad
\bm{\nu}_{k|t}^{r}=\bigl[V_{k|t}^{r},\delta_{k|t}^{r}\bigr]^\top .
\label{eq:mpcc_eu}
\end{align}
The implemented MPCC penalizes the normal contouring error but does not assign
a separate weight to $e_{\ell,k|t}$.  This choice avoids forcing the vehicle
to chase the time stamp of a replanned trajectory; longitudinal timing is
handled through the reference speed term in $\bm{e}_{\nu,k|t}$.

Let
\begin{equation}
Q_c=q_c,\qquad Q_\psi=q_\psi,\qquad
R_\nu=\diag(r_V,r_\delta),\qquad q_{\Delta\delta}>0 .
\label{eq:mpcc_weights}
\end{equation}
For a control sequence
$\mathcal{U}_t=\{\bm{\nu}_{0|t},\ldots,\bm{\nu}_{N_c-1|t}\}$, the
prediction states are induced by~(\ref{eq:mpcc_model}), and the cost is
\begin{equation}
\begin{aligned}
J_t(\mathcal{U}_t)
=&\;q_{\Delta\delta}
\bigl(\delta_{0|t}-\delta_{-1|t}\bigr)^2\\
&+\sum_{k=0}^{N_c-1}
\left[
q_c e_{c,k+1|t}^2
+q_\psi e_{\psi,k+1|t}^2
+\bm{e}_{\nu,k|t}^{\top}R_\nu\bm{e}_{\nu,k|t}
\right].
\end{aligned}
\label{eq:mpcc_cost}
\end{equation}
Equivalently, after substituting
(\ref{eq:mpcc_ec})--(\ref{eq:mpcc_eu}), the stage term is
\begin{equation}
\begin{aligned}
L_{k|t}
=&\;q_c\!
\left[-(x_{k+1|t}-x_{k+1|t}^{r})\sin\theta_{k+1|t}^{r}
+(y_{k+1|t}-y_{k+1|t}^{r})\cos\theta_{k+1|t}^{r}\right]^2\\
&+q_\psi
\left[\operatorname{wrap}(\theta_{k+1|t}-\theta_{k+1|t}^{r})\right]^2
+r_V(V_{k|t}-V_{k|t}^{r})^2
+r_\delta(\delta_{k|t}-\delta_{k|t}^{r})^2 .
\end{aligned}
\label{eq:mpcc_stage_expanded}
\end{equation}
Thus the optimizer is an MPCC because the lateral objective is written in the
moving tangent--normal frame of the planned path, while its decision variables
remain the physically meaningful speed and steering commands.

The nonlinear program solved at each control cycle is
\begin{equation}
\begin{aligned}
\mathcal{U}_t^\star
=\argmin_{\mathcal{U}_t}\quad&
J_t(\mathcal{U}_t)\\
\mathrm{s.t.}\quad&
\bm{\chi}_{0|t}=\bm{0},\\
&\bm{\chi}_{k+1|t}=f_d^{\rm mpcc}
(\bm{\chi}_{k|t},\bm{\nu}_{k|t}),\\
&V_{\min}\le V_{k|t}\le V_{\max},\\
&|\delta_{k|t}|\le\delta_{\max}^{c},\\
&|\Delta V_{k|t}|\le\Delta V_{\max},\\
&|\Delta\delta_{k|t}|\le\dot\delta_{\max}^{c}T_c,\qquad
k=0,\ldots,N_c-1,
\end{aligned}
\label{eq:mpcc_nlp}
\end{equation}
where $f_d^{\rm mpcc}$ denotes the discrete model in
(\ref{eq:mpcc_model})--(\ref{eq:yaw_rate_fit}), and
\begin{equation}
\Delta V_{k|t}=V_{k|t}-V_{k-1|t},\qquad
\Delta\delta_{k|t}=\delta_{k|t}-\delta_{k-1|t}.
\label{eq:mpcc_rate}
\end{equation}
The previous command in~(\ref{eq:mpcc_initial}) defines the $k=0$ increments.
The optimization variables are only the input sequence; the state sequence is
generated recursively by the model.  This condensed form is the reason the
problem remains small despite being nonlinear.

The NLP is solved by a warm-started primal--dual interior-point method.  If
\begin{equation}
\mathcal{U}_{t-1}^\star=
\{(V_{0|t-1}^\star,\delta_{0|t-1}^\star),\ldots,
(V_{N_c-1|t-1}^\star,\delta_{N_c-1|t-1}^\star)\},
\end{equation}
the initial guess at the next cycle is shifted as
\begin{equation}
\mathcal{U}_{t}^{(0)}=
\{(V_{1|t-1}^\star,\delta_{1|t-1}^\star),\ldots,
(V_{N_c-1|t-1}^\star,\delta_{N_c-1|t-1}^\star),
(V_{N_c-1|t-1}^\star,\delta_{N_c-1|t-1}^\star)\}.
\label{eq:mpcc_warm_start}
\end{equation}
The receding-horizon command is
\begin{equation}
V_{\rm cmd}=V_{0|t}^{\star},\qquad
\delta_{\rm cmd}=\delta_{0|t}^{\star},\qquad
\widehat r_{\rm cmd}=\Omega(V_{\rm cmd},\delta_{\rm cmd}).
\label{eq:mpcc_command}
\end{equation}
A solution is used only when the interior-point solver reports convergence
and the objective value is below the admissible tracking threshold; otherwise
the last stabilizing tracking command is retained for the current cycle.

The optimizer output is not sent to the vehicle as an abstract acceleration.
It is converted into the actuator vector
\begin{equation}
\bm{a}_{\rm act}=
\bigl[T_{\rm cmd},\bm{P}_{b,{\rm cmd}},S_{\rm cmd},g_{\rm cmd}\bigr],
\label{eq:actuator_vector}
\end{equation}
where $T_{\rm cmd}$ is throttle opening, $\bm{P}_{b,{\rm cmd}}$ is the
four-wheel brake-pressure command, $S_{\rm cmd}$ is the steering actuator
command, and $g_{\rm cmd}$ is the gear.  The longitudinal force request is
computed from a speed feedforward--feedback law,
\begin{align}
e_{V,t} &= V_{\rm cmd}-V_{\rm meas},
\label{eq:speed_error_ctrl}\\
I_{V,t} &=
\operatorname{clip}\bigl(I_{V,t-1}+T_c e_{V,t},
I_{\min}(e_{V,t}),I_{\max}(e_{V,t})\bigr),
\label{eq:int_speed_ctrl}\\
F_{\rm fb} &=K_p^V e_{V,t}+K_i^V I_{V,t},
\label{eq:fb_force_ctrl}\\
F_{\rm ff} &=c_{d,2}V_{\rm meas}^2+c_{d,1}V_{\rm meas}+c_{d,0}
+m\frac{V_{\rm la}-V_{\rm cmd}}{T_{\rm la}},
\label{eq:ff_force_ctrl}\\
F_x^{\rm req} &=F_{\rm ff}+F_{\rm fb}.
\label{eq:total_force_ctrl}
\end{align}
Here $V_{\rm la}$ is the look-ahead speed sampled on the planned path.  The
engine force available in gear $g$ is
\begin{equation}
F_{\max}(\omega_e,g)=
\frac{i_g i_0}{r_w}
\sum_{j=0}^{5}p_j\omega_e^{5-j},
\label{eq:engine_force}
\end{equation}
where $i_g$ and $i_0$ are the gear and final-drive ratios, $r_w$ is wheel
radius, and $\{p_j\}$ are the fitted engine-torque coefficients.  The
throttle/brake split is
\begin{equation}
T_{\rm raw}=
\begin{cases}
\max\!\left(100F_x^{\rm req}/F_{\max}(\omega_e,g),\,T_0\right),
& F_x^{\rm req}>0,\\
T_0\!\left(1-F_x^{\rm req}/F_0\right),
& F_0<F_x^{\rm req}\le0,\\
0, & F_x^{\rm req}\le F_0,
\end{cases}
\label{eq:throttle_map}
\end{equation}
and
\begin{equation}
P_b=
\begin{cases}
0, & F_x^{\rm req}>F_0,\\
\operatorname{clip}\!\left(
\dfrac{-F_x^{\rm req}r_w\beta_b}{2k_b}\,10^5,\;
P_{b,\min},P_{b,\max}\right),
& F_x^{\rm req}\le F_0,
\end{cases}
\label{eq:brake_map}
\end{equation}
where $T_0$ is the zero-force throttle, $F_0<0$ is the force threshold at
which braking begins, $\beta_b$ is the front brake bias, and $k_b$ converts
brake pressure to wheel torque.  Front and rear pressures are assigned by
\begin{equation}
\bm{P}_{b}^{0}=
\bigl[P_b,\;P_b,\;P_b(1-\beta_b)/\beta_b,\;
P_b(1-\beta_b)/\beta_b\bigr]^\top ,
\label{eq:brake_split}
\end{equation}
and the final brake vector is obtained from the ABS limiter,
\begin{equation}
\bm{P}_{b,{\rm cmd}}=
\mathcal{A}_{\rm ABS}
\bigl(\bm{P}_{b}^{0},\bm{\lambda},\dot{\bm{\lambda}},
|\bm{\alpha}|\bigr).
\label{eq:abs_map}
\end{equation}
The throttle command is clipped by the traction limiter,
\begin{equation}
T_{\rm cmd}=
\operatorname{clip}\!\left(
\min\{T_{\rm raw},T_{\rm tc}(\alpha_r,V_{\rm meas})\},
T_{\min},T_{\max}\right).
\label{eq:tc_map}
\end{equation}

The steering actuator command is a calibrated affine map of the MPCC front
wheel angle,
\begin{equation}
S_{\rm cmd}=
\operatorname{clip}\!\left(
\mathcal{S}_{\delta}\bigl(r_s(\delta_{\rm cmd}+\delta_{\rm off})\bigr),
S_{\min},S_{\max}\right),
\label{eq:steer_actuator_map}
\end{equation}
where $r_s$ is the steering ratio and $\delta_{\rm off}$ is the identified
steering offset.  Gear selection is hysteretic:
\begin{equation}
g_{\rm cmd}=
\begin{cases}
g+1,& \omega_e\ge\omega_{\uparrow,g},\;g<G,\;
F_x^{\rm req}>0,\;\Delta t_g\ge T_{\rm shift},\\
g-1,& \omega_e<\omega_{\downarrow,g},\;g>1,\;
F_x^{\rm req}\le0,\;\Delta t_g\ge T_{\rm shift},\\
1,& g=0,\;V_{\rm cmd}>V_{\min},\\
g,& \text{otherwise}.
\end{cases}
\label{eq:gear_logic}
\end{equation}
Equations~(\ref{eq:mpcc_nlp})--(\ref{eq:gear_logic}) define the complete
control law used in the experiments: the MPCC determines the locally feasible
speed--steering pair, while the calibrated actuator layer realizes that pair
as throttle, brake, steering and gear commands under the vehicle limits.

\input{simulation_results_refined.tex}

% =====================================================================
\section{Conclusion}\label{sec:conclusion}
% =====================================================================
This paper presented a hierarchical tactical--planning--control framework for
wheel-to-wheel autonomous racing at Formula-class speeds.  The central idea is
to restrict learning to the tactical feasible set and to retain auditable
model-based optimization for path planning, speed profiling, and closed-loop
control.  A map-conditioned TQC policy and a hysteretic FSM synthesize the
overtake side, speed cap, and time-varying Frenet corridor; a heading-aware
corridor carver converts the tactical decision into spatial constraints; a
sparse SOCP computes a curvature-regularized path inside this corridor; a
forward--backward sweep enforces the super-elliptic GGV handling envelope; and
an MPCC with terminal progress reward and soft rear-axle friction constraint
executes the reference through the vehicle interface.

The resulting architecture separates tactical intent from physical execution.
This separation is crucial in racing: the learned policy can choose where the
vehicle is allowed to drive, but it cannot directly overwrite tire limits,
actuator bounds, or controller dynamics.  The semi-physical A2RL simulator
experiments validate this design under a virtual-CAN communication loop rather
than an offline trajectory benchmark.  In the short fast-closure scenario, the
system completed the pass $1.13\,\mathrm{s}$ after entering
\textsc{Overtake}, reached $73.10\,\mathrm{m/s}$, and achieved a lateral RMS
tracking error of $0.02\,\mathrm{m}$.  In the longer side-by-side scenario,
the same frozen policy, planner, and controller completed the pass in
$8.63\,\mathrm{s}$ with a lateral RMS error of $0.05\,\mathrm{m}$.  Across
1000 closed-loop semi-physical laps, the stack achieved a $94.7\%$ overtake
success rate without online learning or case-specific retuning.

These results support three conclusions.  First, map-conditioned tactical
learning can resolve the discrete side-selection and commitment problem that
is difficult to encode in a single-vehicle raceline tracker.  Second,
constraining a convex local planner through a learned corridor provides a
transparent interface between data-driven tactical reasoning and
vehicle-dynamics feasibility.  Third, the MPCC layer can execute distinct
tactical corridors with sub-decimeter lateral accuracy when the upstream
planner preserves curvature smoothness and handling-limit speed consistency.
Future work will extend the validation to multi-opponent races, incorporate
probabilistic opponent prediction tubes into the SOCP corridor, and transfer
the semi-physical results to controlled track tests on the Dallara EAV24
platform.

\begin{thebibliography}{38}

% --- Time-optimal trajectory & lap-time ---
\bibitem{christ2021}
F.~Christ, A.~Wischnewski, A.~Heilmeier, and B.~Lohmann, ``Time-optimal trajectory planning for a race car considering variable tyre--road friction coefficients,'' \emph{Veh.\ Syst.\ Dyn.}, vol.~59, no.~4, pp.~588--612, 2021.

\bibitem{heilmeier2020}
A.~Heilmeier, A.~Wischnewski, L.~Hermansdorfer, J.~Betz, M.~Lienkamp, and B.~Lohmann, ``Minimum curvature trajectory planning and control for an autonomous race car,'' \emph{Veh.\ Syst.\ Dyn.}, vol.~58, no.~10, pp.~1497--1527, 2020.

\bibitem{kapania2016}
N.~R.~Kapania, J.~Subosits, and J.~C.~Gerdes, ``A sequential two-step algorithm for fast generation of vehicle racing trajectories,'' \emph{ASME J.\ Dyn.\ Syst.\ Meas.\ Control}, vol.~138, no.~9, art.~091005, 2016.

\bibitem{lipp2014}
T.~Lipp and S.~Boyd, ``Minimum-time speed optimisation over a fixed path,'' \emph{Int.\ J.\ Control}, vol.~87, no.~6, pp.~1297--1311, 2014.

\bibitem{verschueren2014}
R.~Verschueren, S.~De~Bruyne, M.~Zanon, J.~V.~Frasch, and M.~Diehl, ``Towards time-optimal race car driving using nonlinear MPC in real-time,'' in \emph{Proc.\ IEEE Conf.\ Decision Control (CDC)}, 2014, pp.~2505--2510.

\bibitem{rosolia2020lmpc}
U.~Rosolia and F.~Borrelli, ``Learning how to autonomously race a car: A predictive control approach,'' \emph{IEEE Trans.\ Control Syst.\ Technol.}, vol.~28, no.~6, pp.~2713--2719, 2020.

\bibitem{brunner2017}
M.~Brunner, U.~Rosolia, J.~Gonzales, and F.~Borrelli, ``Repetitive learning model predictive control: An autonomous racing example,'' in \emph{Proc.\ IEEE Conf.\ Decision Control (CDC)}, 2017, pp.~2545--2550.

% --- Full-stack racing platforms ---
\bibitem{kabzan2020amzracing}
J.~Kabzan, M.~de~la~Iglesia~Valls, V.~Reijgwart \emph{et al.}, ``AMZ driverless: The full autonomous racing system,'' \emph{J.\ Field Robot.}, vol.~37, no.~7, pp.~1267--1294, 2020.

\bibitem{wischnewski2022indy}
A.~Wischnewski, M.~Geisslinger, J.~Betz \emph{et al.}, ``Indy autonomous challenge---Autonomous race cars at the handling limits,'' in \emph{Proc.\ FISITA World Automot.\ Congr.}, 2022, pp.~1--14.

\bibitem{betz2023autonomous}
J.~Betz, T.~Betz, F.~Fent \emph{et al.}, ``TUM autonomous motorsport: An autonomous racing software for the Indy autonomous challenge,'' \emph{J.\ Field Robot.}, vol.~40, no.~4, pp.~783--809, 2023.

\bibitem{jung2023iac}
C.~Jung, A.~Finazzi, H.~Seong \emph{et al.}, ``An autonomous system for head-to-head race: Design, implementation and analysis,'' \emph{IEEE Trans.\ Intell.\ Transp.\ Syst.}, vol.~24, no.~10, pp.~10470--10486, 2023.

\bibitem{raji2024kth}
A.~Raji, A.~Liniger, A.~Giove \emph{et al.}, ``A tricycle model to accurately control an autonomous racecar with locked differentials,'' \emph{IEEE Robot.\ Autom.\ Lett.}, vol.~9, no.~3, pp.~2499--2506, 2024.

% --- MPC at handling limit / MPCC / GP ---
\bibitem{kabzan2019gpmpc}
J.~Kabzan, L.~Hewing, A.~Liniger, and M.~N.~Zeilinger, ``Learning-based model predictive control for autonomous racing,'' \emph{IEEE Robot.\ Autom.\ Lett.}, vol.~4, no.~4, pp.~3363--3370, 2019.

\bibitem{hewing2020}
L.~Hewing, K.~P.~Wabersich, M.~Menner, and M.~N.~Zeilinger, ``Learning-based model predictive control: Toward safe learning in control,'' \emph{Annu.\ Rev.\ Control Robot.\ Auton.\ Syst.}, vol.~3, pp.~269--296, 2020.

\bibitem{liniger2015}
A.~Liniger, A.~Domahidi, and M.~Morari, ``Optimization-based autonomous racing of 1:43 scale RC cars,'' \emph{Optim.\ Control Appl.\ Methods}, vol.~36, no.~5, pp.~628--647, 2015.

\bibitem{liniger2019rss}
A.~Liniger and J.~Lygeros, ``A noncooperative game approach to autonomous racing,'' \emph{IEEE Trans.\ Control Syst.\ Technol.}, vol.~28, no.~3, pp.~884--897, 2020.

\bibitem{romero2022}
A.~Romero, S.~Sun, P.~Foehn, and D.~Scaramuzza, ``Model predictive contouring control for time-optimal quadrotor flight,'' \emph{IEEE Trans.\ Robot.}, vol.~38, no.~6, pp.~3340--3356, 2022.

\bibitem{novi2020}
T.~Novi, A.~Liniger, R.~Capitani, and C.~Annicchiarico, ``Real-time control for at-limit handling driving on a predefined path,'' \emph{Veh.\ Syst.\ Dyn.}, vol.~58, no.~7, pp.~1007--1036, 2020.

\bibitem{verschueren2022acados}
R.~Verschueren, G.~Frison, D.~Kouzoupis \emph{et al.}, ``acados---A modular open-source framework for fast embedded optimal control,'' \emph{Math.\ Program.\ Comput.}, vol.~14, pp.~147--183, 2022.

\bibitem{domahidi2014forces}
A.~Domahidi and J.~Jerez, ``FORCES Professional,'' Embotech AG, 2014--2024. [Online]. Available: \url{https://embotech.com/forces-pro}.

% --- Game-theoretic & multi-agent ---
\bibitem{wang2021iterative}
M.~Wang, Z.~Wang, J.~Talbot, J.~C.~Gerdes, and M.~Schwager, ``Game-theoretic planning for self-driving cars in multivehicle competitive scenarios,'' \emph{IEEE Trans.\ Robot.}, vol.~37, no.~4, pp.~1313--1325, 2021.

\bibitem{schwarting2021}
W.~Schwarting, T.~Seyde, I.~Gilitschenski, L.~Liebenwein, R.~Sander, S.~Karaman, and D.~Rus, ``Deep latent competition: Learning to race using visual control policies in latent space,'' in \emph{Proc.\ Conf.\ Robot Learn.\ (CoRL)}, 2021, pp.~1855--1870.

\bibitem{fridovich2020ilqgames}
D.~Fridovich-Keil, E.~Ratner, L.~Peters, A.~D.~Dragan, and C.~J.~Tomlin, ``Efficient iterative linear-quadratic approximations for nonlinear multi-player general-sum differential games,'' in \emph{Proc.\ IEEE Int.\ Conf.\ Robot.\ Autom.\ (ICRA)}, 2020, pp.~1475--1481.

\bibitem{laine2023}
F.~Laine, D.~Fridovich-Keil, C.-Y.~Chiu, and C.~J.~Tomlin, ``The computation of approximate generalized feedback Nash equilibria,'' \emph{SIAM J.\ Optim.}, vol.~33, no.~1, pp.~294--318, 2023.

\bibitem{spica2020}
R.~Spica, E.~Cristofalo, Z.~Wang, E.~Montijano, and M.~Schwager, ``A real-time game theoretic planner for autonomous two-player drone racing,'' \emph{IEEE Trans.\ Robot.}, vol.~36, no.~5, pp.~1389--1403, 2020.

\bibitem{liniger2020stackelberg}
A.~Liniger and J.~Lygeros, ``Real-time control for autonomous racing based on viability theory,'' \emph{IEEE Trans.\ Control Syst.\ Technol.}, vol.~27, no.~2, pp.~464--478, 2019.

\bibitem{williams2018best}
G.~Williams, P.~Drews, B.~Goldfain, J.~M.~Rehg, and E.~A.~Theodorou, ``Information-theoretic model predictive control: Theory and applications to autonomous driving,'' \emph{IEEE Trans.\ Robot.}, vol.~34, no.~6, pp.~1603--1622, 2018.

\bibitem{wurman2022gtsophy}
P.~R.~Wurman, S.~Barrett, K.~Kawamoto \emph{et al.}, ``Outracing champion Gran Turismo drivers with deep reinforcement learning,'' \emph{Nature}, vol.~602, no.~7896, pp.~223--228, 2022.

\bibitem{jia2023bayesianrace}
H.~Jia, Y.~Wu, X.~Yang, J.~Pan, and L.~Liu, ``Bayesian intent inference for head-to-head autonomous racing,'' \emph{IEEE Trans.\ Intell.\ Veh.}, vol.~8, no.~7, pp.~3851--3863, 2023.

% --- RL on real hardware / safe RL ---
\bibitem{fuchs2021superhuman}
F.~Fuchs, Y.~Song, E.~Kaufmann, D.~Scaramuzza, and P.~Duerr, ``Super-human performance in Gran Turismo Sport using deep reinforcement learning,'' \emph{IEEE Robot.\ Autom.\ Lett.}, vol.~6, no.~3, pp.~4257--4264, 2021.

\bibitem{song2021autonomous}
Y.~Song, H.~Lin, E.~Kaufmann, P.~Duerr, and D.~Scaramuzza, ``Autonomous overtaking in Gran Turismo Sport using curriculum reinforcement learning,'' in \emph{Proc.\ IEEE Int.\ Conf.\ Robot.\ Autom.\ (ICRA)}, 2021, pp.~9403--9409.

\bibitem{evans2023safelearning}
B.~D.~Evans, H.~A.~Engelbrecht, and H.~W.~Jordaan, ``Safe reinforcement learning for high-speed autonomous racing,'' \emph{Cogn.\ Robot.}, vol.~3, pp.~107--126, 2023.

\bibitem{brunnbauer2022learning}
A.~Brunnbauer, L.~Berducci, A.~Brandst\"atter \emph{et al.}, ``Latent imagination facilitates zero-shot transfer in autonomous racing,'' in \emph{Proc.\ IEEE Int.\ Conf.\ Robot.\ Autom.\ (ICRA)}, 2022, pp.~7513--7520.

\bibitem{he2023ralrace}
S.~He, J.~Zeng, B.~Zhang, and K.~Sreenath, ``Rule-based safety-critical control design using control barrier functions with application to autonomous lane change,'' \emph{IEEE Trans.\ Control Syst.\ Technol.}, vol.~31, no.~6, pp.~2701--2715, 2023.

\bibitem{ye2021mcts}
Y.~Ye, X.~Zhang, and J.~Sun, ``Automated vehicle's behavior decision making using deep reinforcement learning and high-fidelity simulation environment,'' \emph{Transp.\ Res.\ Part C}, vol.~107, pp.~155--170, 2019.

\bibitem{sunberg2017pomdp}
Z.~N.~Sunberg, C.~J.~Ho, and M.~J.~Kochenderfer, ``The value of inferring the internal state of traffic participants for autonomous freeway driving,'' in \emph{Proc.\ Amer.\ Control Conf.\ (ACC)}, 2017, pp.~3004--3010.

% --- Surveys & corridor / safe set ---
\bibitem{paden2016survey}
B.~Paden, M.~\v{C}\'ap, S.~Z.~Yong, D.~Yershov, and E.~Frazzoli, ``A survey of motion planning and control techniques for self-driving urban vehicles,'' \emph{IEEE Trans.\ Intell.\ Veh.}, vol.~1, no.~1, pp.~33--55, 2016.

\bibitem{gonzalez2016review}
D.~Gonz\'alez, J.~P\'erez, V.~Milan\'es, and F.~Nashashibi, ``A review of motion planning techniques for automated vehicles,'' \emph{IEEE Trans.\ Intell.\ Transp.\ Syst.}, vol.~17, no.~4, pp.~1135--1145, 2016.

\bibitem{betz2022software}
J.~Betz, H.~Zheng, A.~Liniger \emph{et al.}, ``Autonomous vehicles on the edge: A survey on autonomous vehicle racing,'' \emph{IEEE Open J.\ Intell.\ Transp.\ Syst.}, vol.~3, pp.~458--488, 2022.

\bibitem{betz2022survey}
J.~Betz, A.~Wischnewski, A.~Heilmeier \emph{et al.}, ``A software architecture for the dynamic path planning of an autonomous racecar at the limits of handling,'' in \emph{Proc.\ IEEE Int.\ Conf.\ Connect.\ Veh.\ Expo (ICCVE)}, 2019, pp.~1--8.

\bibitem{kaufmann2023champion}
E.~Kaufmann, L.~Bauersfeld, A.~Loquercio \emph{et al.}, ``Champion-level drone racing using deep reinforcement learning,'' \emph{Nature}, vol.~620, no.~7976, pp.~982--987, 2023.

\bibitem{ziegler2014baidu}
J.~Ziegler, P.~Bender, M.~Schreiber \emph{et al.}, ``Making Bertha drive---An autonomous journey on a historic route,'' \emph{IEEE Intell.\ Transp.\ Syst.\ Mag.}, vol.~6, no.~2, pp.~8--20, 2014.

\bibitem{werling2010frenet}
M.~Werling, J.~Ziegler, S.~Kammel, and S.~Thrun, ``Optimal trajectory generation for dynamic street scenarios in a Fr\'enet frame,'' in \emph{Proc.\ IEEE Int.\ Conf.\ Robot.\ Autom.\ (ICRA)}, 2010, pp.~987--993.

\bibitem{rosolia2018sample}
U.~Rosolia and F.~Borrelli, ``Learning model predictive control for iterative tasks. A data-driven control framework,'' \emph{IEEE Trans.\ Autom.\ Control}, vol.~63, no.~7, pp.~1883--1896, 2018.

\bibitem{brunke2022safe}
L.~Brunke, M.~Greeff, A.~W.~Hall \emph{et al.}, ``Safe learning in robotics: From learning-based control to safe reinforcement learning,'' \emph{Annu.\ Rev.\ Control Robot.\ Auton.\ Syst.}, vol.~5, pp.~411--444, 2022.

\bibitem{lin2024racellm}
H.~Lin, Y.~Song, and D.~Scaramuzza, ``Reaching the limit in autonomous racing: Optimal control versus reinforcement learning,'' \emph{Sci.\ Robot.}, vol.~8, no.~82, art.~eadg1462, 2023.

\bibitem{remonda2024}
A.~Remonda, E.~Veas, and G.~Luzhnica, ``Comparing the performance of reinforcement learning algorithms for autonomous racing,'' \emph{Eng.\ Appl.\ Artif.\ Intell.}, vol.~118, art.~105692, 2023.

\end{thebibliography}

\end{document}
